---
title: "EDA"
author: "Thomas FitzGerald"
date: "2024-10-29"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)

library(tidyverse)
library(dplyr)


library(ggplot2)
library(ggridges)
library(plotly)

library(stats)
library(multcomp)
library(emmeans)
library(lsmeans)
library(stringr)
library(lubridate)
```

```{r}
variable.unit.converter <- function(df,df.datadict,var.name,unit.type.new){
  #Retrieve information from data dictionary
  var1 <- df[var.name]
  var.type <- subset(df.datadict,name==var.name)$measurement_type
  unit.type <- subset(df.datadict,name==var.name)$measurement_unit

  #Check   
  #if (unit.type==unit.type.new) {return(var1)}
  
  if(var.type=='hourly_wage') {
    hours.var <- df[paste(var.name,'_hours',sep='')]
    
    if (unit.type.new == 'minute') {var1 <- var1/60} else 
      if (unit.type.new == 'hour') {var1 <- var1} else 
      if (unit.type.new == 'day') {var1 <- var1 * hours.var / 5} else 
      if (unit.type.new == 'week') {var1 <- var1 * hours.var} else 
      if (unit.type.new == 'month') {var1 <- var1 * hours.var * 4.3} else 
      if (unit.type.new == 'year') {var1 <- var1 * hours.var * 52} else 
      {var1 <- var1}
  }
  if (var.type == 'weight') {
    if (unit.type.new == 'ounce') {var1 <- var1/16} else
      if (unit.type.new == 'pound') {var1 <- var1} else
      {var1 <- var1}
  }
  if (var.type == 'volume') {
    if (unit.type.new == 'ounce') {var1 <- var1/128} else
      if (unit.type.new == 'gallon') {var1 <- var1} else
      {var1 <- var1}
  }
  return(var1)
}
```

```{r}
dir <- 'C:\\Users\\Thoma\\OneDrive\\Desktop\\portoflio_project_1\\data'
df <- read.csv(paste(dir,'main_data_file.csv',sep='\\'))
df.datadict <- read.csv(paste(dir,'data_dictionary.csv',sep='\\'))
df
colnames(df)
df.datadict
```

```{r}
df %>% dplyr::select(colnames(df)[grep("hours",colnames(df),perl=TRUE)])
df %>% dplyr::select(colnames(df)[grep("hourly",colnames(df),perl=TRUE)])
c('manufacturing_general')
```

```{r}
plot_ly(data=df,x=~year,mode='lines+markers') %>%
  add_trace(y = ~manufacturing_general) %>%
  add_trace(y = ~retail) %>%
  add_trace(y = ~mining_logging_wages)
```

Dictionary of {display_name:var_name}
```{r}

base_var <- df$manufacturing_general
var1 <- df$gallon_of_milk
var2 <- df$pound_of_ground_beef
prop1 <- var1/base_var
prop2 <- var2/base_var

df.display <- as.data.frame(cbind(df$year,prop1,prop2))
colnames(df.display) <- c('year','Retail','Mining')


plot_ly(data=df.display,x=~year,mode='lines+markers') %>%
  add_trace(y = ~Retail) %>%
  add_trace(y = ~Mining)
```
Currently, if you do wage base, product var1/var2, you get a proportion.  Instead, we want to display it in a relevent unit.

So, currently, 1 loaf of bread might be .14 hours.  Instead, we would want this listed as 8.4 minutes.

So, you select "manufacturing labor" as your base unit.  Then, you can select minutes, hours, days, weeks, months, or years.

  *Later, we could add an intelligent scaling option, which would pick a unit that is as close as possible to 1-100 units.  So, bread would scale to minutes, since it runs 8-~20 minutes.




```{r}
df <- df
df.datadict <- df.datadict
var.name <- 'gallon_of_gasoline'
unit.type.new <- 'gallon'

variable.unit.converter <- function(df,df.datadict,var.name,unit.type.new){
  #Retrieve information from data dictionary
  var1 <- df[var.name]
  var.type <- subset(df.datadict,name==var.name)$measurement_type
  unit.type <- subset(df.datadict,name==var.name)$measurement_unit

  #Check   
  #if (unit.type==unit.type.new) {return(var1)}
  
  if(var.type=='hourly_wage') {
    hours.var <- df[paste(var.name,'_hours',sep='')]
    
    if (unit.type.new == 'minute') {var1 <- var1/60} else 
      if (unit.type.new == 'hour') {var1 <- var1} else 
      if (unit.type.new == 'day') {var1 <- var1 * hours.var / 5} else 
      if (unit.type.new == 'week') {var1 <- var1 * hours.var} else 
      if (unit.type.new == 'month') {var1 <- var1 * hours.var * 4.3} else 
      if (unit.type.new == 'year') {var1 <- var1 * hours.var * 52} else 
      {var1 <- var1}
  }
  if (var.type == 'weight') {
    if (unit.type.new == 'ounce') {var1 <- var1/16} else
      if (unit.type.new == 'pound') {var1 <- var1} else
      {var1 <- var1}
  }
  if (var.type == 'volume') {
    if (unit.type.new == 'ounce') {var1 <- var1/128} else
      if (unit.type.new == 'gallon') {var1 <- var1} else
      {var1 <- var1}
  }
  return(var1)
}
df$retail_hours[125]
variable.unit.converter(df,df.datadict,'retail','minute')[125,1]
variable.unit.converter(df,df.datadict,'retail','hour')[125,1]
variable.unit.converter(df,df.datadict,'retail','day')[125,1]
variable.unit.converter(df,df.datadict,'retail','week')[125,1]
variable.unit.converter(df,df.datadict,'retail','month')[125,1]
variable.unit.converter(df,df.datadict,'retail','year')[125,1]
sort(colnames(df))
```

```{r}
#v1
base_var.name <- 'flour'
base_var.unit <- 'pound'
var1.name <- 'retail'
var1.unit <- 'hour'
var2.name <- 'financial'
var2.unit <- 'hour'

base_var <- variable.unit.converter(df,df.datadict,base_var.name,base_var.unit)

var1 <- variable.unit.converter(df,df.datadict,var1.name,var1.unit)
var2 <- variable.unit.converter(df,df.datadict,var2.name,var2.unit)
prop1 <- var1/base_var
prop2 <- var2/base_var

df.display <- as.data.frame(cbind(df$year,prop1,prop2))
colnames(df.display) <- c('year',colnames(var1),colnames(var2))

plot_ly(data=df.display %>% pivot_longer(cols=c(colnames(var1),colnames(var2)),names_to='names',values_to='values'),
        x=~year,mode='lines+markers', split= ~names) %>%
  add_trace(y = ~values) %>%
  layout(legend=list(title=list(text='Variables')),
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'Gallons of Gasoline per Hour'))
```

```{r}
#v2
base_var.name <- 'ground_beef'
base_var.unit <- 'pound'
var.names <- c('retail','financial','mining_logging_wages')
var.units <- c('hour','hour','hour')

base_var_scatterplot <- function(df,df.datadict,base_var.name,base_var.unit,var.names,var.units) {
  base_var <- variable.unit.converter(df,df.datadict,base_var.name,base_var.unit)
  df.display <- as.data.frame(df$year)
  
  for (i in 1:length(var.names)) {
    var <- variable.unit.converter(df,df.datadict,var.names[i],var.units[i])
    prop <- var/base_var
    df.display <- as.data.frame(cbind(df.display,prop))
  }
  
  colnames(df.display) <- c('year',var.names)
  df.display <- df.display %>% pivot_longer(cols=var.names)
    
  plot_ly(data=df.display,
          x=~year,mode='lines+markers', split= ~name) %>%
    add_trace(y = ~value) %>%
    layout(legend=list(title=list(text='Variables')),
           xaxis = list(title = 'Year'),
           yaxis = list(title = 'Gallons of Gasoline per Hour'))
}

base_var_scatterplot(df,df.datadict,base_var.name,base_var.unit,var.names,var.units)
```
```{r}

#v2
base_var.name <- 'ground_beef'
base_var.unit <- 'pound'
var.names <- c('retail','financial','mining_logging_wages')
var.units <- c('unit','unit','unit')

create_df_display <- function(df,df.datadict,base_var.name,base_var.unit,var.names,var.units) {
  base_var <- variable.unit.converter(df,df.datadict,base_var.name,base_var.unit)
  df.display <- as.data.frame(df$year)
  
  for (i in 1:length(var.names)) {
    var <- variable.unit.converter(df,df.datadict,var.names[i],var.units[i])
    prop <- var/base_var
    df.display <- as.data.frame(cbind(df.display,prop))
  }
  
  colnames(df.display) <- c('year',var.names)
  df.display <- df.display %>% pivot_longer(cols=var.names)
  return(df.display)
}

create_df_display(df,df.datadict,base_var.name,base_var.unit,var.names,var.units)
```

df$manufacturing_durable_goods_hourly

Base_Variable
  Dollars
  Wages
    minute
    hour*
    day
    week
    month
    year
  PhysicalGoods

  Weight
    Ounces
    Pounds*
  Volume
    Ounces
    Gallon

hourly
weekly
annually

minute = hourly * 60
hour = hourly
day = weekly / 5
weekly = hourly * hours
monthly = weekly * 4.345
yearly = weekly * 52

ounces = pounds * 16
ounces = gallon * 128

var.name <- 'retail'
subset(df.datadict,name==var.name)$measurement_type

df[paste(var.name,'_hours',sep='')]
df[paste(var.name,'_hourly',sep='')]

```

```{r}
    # myplot <- plot_ly(data=df,
    #                   x=year,
    #                   mode = 'markers') %>%
      # layout(
      #   title = paste(input.r$region.1.r,': ',input.r$cod_1.r,sep=""),
      #   # xaxis = list(title="Month",
      #   #              range=c(1,52),
      #   #              showgrid = TRUE,
      #   #              ticktext = c("Jan","Feb","Mar","Apr","May","June","July","Aug","Sep","Oct","Nov","Dec"),
      #   #              tickvals = c(1,5,9,14,18,22,27,31,35,40,45,49)
      #   #              ),
      #   yaxis = list(title="Mortality",range=c(0,y.top)),
      #   showlegend = TRUE,
      #   legend = list(
      #     title=list(text="Year", font=list(color="black", size=15)),
      #     orientation = 'h',
      #     xanchor = "center",
      #     x = .5,
      #     y = -.2
        )
      )


```




