---
title: "server_test"
author: "Thomas FitzGerald"
date: "2025-03-26"
output: html_document
---

Broad Outline:
Libraries
Import Data
Initial Data Modification

Color Palette
Define Drop-down lists

shinyServer
  Reactive Variables
  observeEvents() functions
  Selectors


```{r setup, include=FALSE}
library(shiny)
library(shinyWidgets)
library(shinydashboard)
library(readr)
library(DT)
library(plotly)
library(ggplot2)
library(tidyverse)
library(stringr)
library(lubridate)
library(RColorBrewer)
#library(usmap)
```

```{r}
#Main Value Data
dir <- 'C:\\Users\\Thoma\\OneDrive\\Desktop\\portoflio_project_1\\data'
df.raw <- read.csv(paste(dir,'main_data_file.csv',sep='\\'))
df.orig <- df.raw #df.raw used for table display.

#Data Dictionary
df.datadict <- read.csv(paste(dir,'data_dictionary.csv',sep='\\'))
```

This section defines drop-down lists (Variable names, etc.)
```{r}
#Set Drop-Down lists
variables.list <- as.data.frame(colnames(df.orig))
colnames(variables.list)<-c("vars")
variables.list <- variables.list %>% 
  subset(!vars %in% c('X','year')) %>%
  filter(!grepl('hours', vars))
variables.list
```

```{r}
#Define Color Palette for Graphs
palette.main<-brewer.pal(n = 8, name = "Spectral")
ty.color<-palette.main[7]
ly.color<-palette.main[3]
lly.color<-palette.main[1]
```

```{r}
#Set column names for drop-down.
friendly_disease_list<-c(
  "All_Causes",
  "All_Natural_Causes",
  "Septicemia",
  "Cancer",
  "Diabetes",
  "Alzheimers",
  "Influenza_and_Pneumonia",
  "Chronic_Lowe_Respiratory_Diseases",
  "Other_Respiratory_Diseases",
  "Nephritis(Kidney_Problems)",
  "Not_Elsewhere_Classified",
  "Heart_Diseases",
  "Cerebrovascular(Stroke,Aneurysm)",
  "Covid_19_Multiple_COD",
  "Covid_19_Underlying_COD"
)
```

Below is the reactive (input.r) variables.  Comments are from butchered version.
```{r}
  input.r <- reactiveValues(
    df.display = df.orig,
    base_var.name = 'ground_beef',
    base_var.unit = 'pound',
    var.names = c('retail','financial'),
    var.units = c('hour','hour')
  )
    # df.ytd.1 = subset(df.main.1,MMWR.Year==update.year),
    # df.ty.1 = subset(df.main.1,MMWR.Year==last.year&ty_flag==TRUE),
    # df.ly.1 = subset(df.main.1,MMWR.Year==last.year),
    # df.lly.1 = subset(df.main.1,MMWR.Year==last.last.year),
    # df.ytd.2 = subset(df.main.2,MMWR.Year==update.year),
    # df.ty.2 = subset(df.main.2,MMWR.Year==last.year&ty_flag==TRUE),
    # df.ly.2 = subset(df.main.2,MMWR.Year==last.year),
    # df.lly.2 = subset(df.main.2,MMWR.Year==last.last.year),
    # current.year.total.1 = 1,
    # last.year.total.1 = 2,
    # last.last.year.total.1 = 3,
    # current.year.total.2 = 4,
    # last.year.total.2 = 5,
    # last.last.year.total.2 = 6,
    # region.1.r = "United States",
    # region.2.r = "United States",
    # cod_1.r = "Covid_19_Underlying_COD",
    # cod_2.r = "Covid_19_Multiple_COD",
    # current.year.total.1 = sum(subset(df.main.1,MMWR.Year==update.year)["disp"],na.rm=TRUE),
    # #Map
    # df.map = df.map,
    # df.map.active = subset(df.map,MMWR.Year==last.year & MMWR.Week==1),
    # map.year.r = last.year,
    # map.week.r = 1,
    # # last.year.total.1 = sum(df.ly.1["disp"],na.rm=TRUE),
    # # last.last.year.total.1 = sum(df.lly.1["disp"],na.rm=TRUE),
    # # current.year.total.2 = sum(df.ytd.1["disp"],na.rm=TRUE),
    # # last.year.total.2 = sum(df.ly.1["disp"],na.rm=TRUE),
    # # last.last.year.total.2 = sum(df.lly.1["disp"],na.rm=TRUE),
    # view.r = c(TRUE,FALSE,FALSE,TRUE,FALSE,FALSE), #Years to display: first 3 plot 1, second 3 plot 2.  TY/LY/LLY.
```
```{r}
# #Set main dataframes to filter from
df.display <- df.orig
```

Below is for server side observed elements.
regionSelector1 -> baseVarSelector
regionSelector2 -> REMOVED

Old
```{r}
  #Top Region Selector
  output$regionSelector <- renderUI({
    selectInput(
      inputId = "myRegion1",
      label = "Select a State, or 'United States' to view all:",
      choices = region.list
    )
  })

  #Reacts to UI element 'regionSelector1'
  observeEvent(input$myRegion1, {
    input.r$region.1.r <- input$myRegion1
    print(paste('Region 1 changed, active region is now:',input.r$region.1.r))
    input.r$df.ytd.1 <-   subset(input.r$df.main.1,Jurisdiction.of.Occurrence==input.r$region.1.r & MMWR.Year==update.year)
    input.r$df.ty.1 <-    subset(input.r$df.main.1,Jurisdiction.of.Occurrence==input.r$region.1.r & MMWR.Year==last.year&ty_flag==TRUE)
    input.r$df.ly.1 <-    subset(input.r$df.main.1,Jurisdiction.of.Occurrence==input.r$region.1.r & MMWR.Year==last.year)
    input.r$df.lly.1 <-   subset(input.r$df.main.1,Jurisdiction.of.Occurrence==input.r$region.1.r & MMWR.Year==last.last.year )
  }) 
```

New
```{r}
  #Variable Selector

  #Reacts to UI element 'baseVarSelector'
  observeEvent(input$baseVar1, {
    input.r$base.var.1.r <- input$baseVar1
    print(paste('Base Variables changed, active base variable is now:',input.r$region.1.r))
  }) 
```