---
title: "server_test"
author: "Thomas FitzGerald"
date: "2025-03-26"
output: html_document
---

Broad Outline:
Libraries
Import Data
Initial Data Modification

Color Palette
Define Drop-down lists

shinyServer
  Reactive Variables
  observeEvents() functions
  Selectors


```{r setup, include=FALSE}
library(shiny)
library(shinyWidgets)
library(shinydashboard)
library(readr)
library(DT)
library(plotly)
library(ggplot2)
library(tidyverse)
library(stringr)
library(lubridate)
library(RColorBrewer)
#library(usmap)
```

```{r}
#Functions######################################################################
variable.unit.options <- function(var.name,df.datadict){
  unit.type <- subset(df.datadict, var_name==var.name)$measurement_type
  #print(unit.type)
  if(length(unit.type)==0 || unit.type=='None'){return('none')}
  if(unit.type=='currency'){output <- 'unit'}
  if(unit.type=='hourly_wage'){output <- c('hour','minute','day','week','month','year')}
  if(unit.type=='weight'){output <- c('pound','ounce')}
  if(unit.type=='volume'){output <- c('gallon','ounce')}
  if(unit.type=='other'){output <- 'unit'}
  return(output)
}

variable.unit.converter <- function(df,df.datadict,var.name,unit.type.new){
  #Retrieve information from data dictionary
  var1 <- df[var.name]
  var.type <- subset(df.datadict,var_name==var.name)$measurement_type
  unit.type <- subset(df.datadict,var_name==var.name)$measurement_unit
  
  if(var.type=='hourly_wage') {
    hours.var <- df[paste(var.name,'_hours',sep='')]
    
    if (unit.type.new == 'minute') {var1 <- var1/60} else 
      if (unit.type.new == 'hour') {var1 <- var1} else 
        if (unit.type.new == 'day') {var1 <- var1 * hours.var / 5} else 
          if (unit.type.new == 'week') {var1 <- var1 * hours.var} else 
            if (unit.type.new == 'month') {var1 <- var1 * hours.var * 4.3} else 
              if (unit.type.new == 'year') {var1 <- var1 * hours.var * 52} else 
              {var1 <- var1}
  }
  if (var.type == 'weight') {
    if (unit.type.new == 'ounce') {var1 <- var1/16} else
      if (unit.type.new == 'pound') {var1 <- var1} else
      {var1 <- var1}
  }
  if (var.type == 'volume') {
    if (unit.type.new == 'ounce') {var1 <- var1/128} else
      if (unit.type.new == 'gallon') {var1 <- var1} else
      {var1 <- var1}
  }
  return(var1)
}

create_df_display <- function(df,df.datadict,base_var.name,base_var.unit,var.names,var.units) {
  base_var <- variable.unit.converter(df,df.datadict,base_var.name,base_var.unit)
  df.display <- as.data.frame(df$year)
  
  for (i in 1:length(var.names)) {
    var <- variable.unit.converter(df,df.datadict,var.names[i],var.units[i])
    prop <- var/base_var
    df.display <- as.data.frame(cbind(df.display,prop))
  }
  
  colnames(df.display) <- c('year',var.names)
  df.display <- df.display %>% pivot_longer(cols=var.names)
  return(df.display)
}

non_max_na_filter <- function(df){
  #Returns df as only rows with non-max NA.  So, if the most NA columns in any row is 10, removes all rows with 10 NAs.
  return(df[which(rowSums(is.na(df)) %in% rowSums(is.na(df))[rowSums(is.na(df))!=max(rowSums(is.na(df)))]),])
}

base_var_scatterplot <- function(df,df.datadict,base_var.name,base_var.unit,var.names,var.units) {
  #print(paste('base_var_scatterplot:', var.names,var.units,sep=' '))
  var.1 <- var.names[1]
  var.2 <- var.names[2]
  var.3 <- var.names[3]
  print(paste('single:', var.names[3],sep=' '))
    var.names.raw <- var.names
  var.names <- var.names[var.names!='None']
  #print(paste('base_var_scatterplot:', var.names,var.units,sep=' '))
  base_var <- variable.unit.converter(df,df.datadict,base_var.name,base_var.unit)

  df.display <- as.data.frame(df$year)
    
  for (i in 1:length(var.names)) {
    var <- variable.unit.converter(df,df.datadict,var.names[i],var.units[i])
    prop <- var/base_var
    df.display <- as.data.frame(cbind(df.display,prop))
  }
  
  colnames(df.display) <- c('year',var.names)
  df.display <- df.display %>% pivot_longer(cols=var.names)
  
  myplot <- suppressMessages(plotly_build(plot_ly(data=df.display,
          x=~year,mode='lines+markers', split= ~name, color = ~name) %>%
    add_trace(y = ~value) %>%
    layout(legend=list(title=list(text='Variables')),
           xaxis = list(title = 'Year'),
           yaxis = list(title = paste(base_var.name, paste('(', base_var.unit,')',sep=''),sep=' ')))))
  return(myplot)
}

```

Import & Formatting
```{r}
#Data###########################################################################
#Main Value Data
dir <- 'C:\\Users\\Thoma\\OneDrive\\Desktop\\portoflio_project_1\\data'
df.raw <- read.csv(paste(dir,'main_data_file.csv',sep='\\'))
df.orig <- non_max_na_filter(df.raw) #df.raw used for table display.

#Data Dictionary
df.datadict <- read.csv(paste(dir,'data_dictionary.csv',sep='\\'))

#Define Color Palette for Graphs
palette.main<-brewer.pal(n = 8, name = "Spectral")
ty.color<-palette.main[7]
ly.color<-palette.main[3]
lly.color<-palette.main[1]

# #Set main dataframes to filter from
df.main.1 <- df.orig

#Set Drop-Down lists
variables.list <- as.data.frame(colnames(df.orig))
#units.list <- c('unit','minute','hour','day','week','month','year','ounce','pound','oun')
colnames(variables.list)<-c("All")
variables.list <- variables.list %>% 
  subset(!All %in% c('X','year')) %>%
  filter(!grepl('hours', All))

var.categories <- df.datadict %>% dplyr::select(var_category,var_name)
commodities.list <- variables.list %>% 
  subset(All %in% subset(var.categories,var_category=='Commodity')$var_name) %>%
  rename(Commodities = All)
employment.list <- variables.list %>% 
  subset(All %in% subset(var.categories,var_category=='Employment')$var_name) %>%
  rename(Employment = All)
currency.list <- variables.list %>% 
  subset(All %in% subset(var.categories,var_category=='Currency')$var_name) %>%
  rename(Currency = All)
```
datadict adjustments
```{r}
df.datadict$display_description <-  paste("Description: ",df.datadict$var_Description," \n",
                                          "Default Unit: ",df.datadict$measurement_unit,"\n",
                                          "Source: ",df.datadict$Source,"\n",
                                          "Source Description: ",df.datadict$source_description,"\n",
                                          sep='')

```


Reactive Values Stand-in (Will need input.r$ in front in server)
```{r}
    base_var.name = 'dollar'
    base_var.unit = 'unit'
    var.name.1.r = 'flour'
    var.unit.1.r = 'pound'
    var.name.2.r = 'milk'
    var.unit.2.r = 'gallon'
    var.name.3.r = 'None'
    var.unit.3.r = 'none'
    
    var.names <- c(var.name.1.r,var.name.2.r,var.name.3.r)
    var.units <- c(var.unit.1.r,var.unit.2.r,var.unit.3.r)
```

Plot
```{r}
    myplot <- base_var_scatterplot(df.orig,
                                   df.datadict,
                                   base_var.name.1.r,base_var.unit.1.r,
                                   c(var.name.1.r,var.name.2.r,var.name.3.r),
                                   c(var.unit.1.r,var.unit.2.r,var.unit.3.r))

```
```{r}
df <- df.orig
base_var_scatterplot <- function(df,df.datadict,base_var.name,base_var.unit,var.names,var.units) {
  #print(paste('base_var_scatterplot:', var.names,var.units,sep=' '))
  var.1 <- var.names[1]
  var.2 <- var.names[2]
  var.3 <- var.names[3]
  print(paste('single:', var.names[3],sep=' '))
  var.names.raw <- var.names
  var.names <- var.names[var.names!='None']
  #print(paste('base_var_scatterplot:', var.names,var.units,sep=' '))
  base_var <- variable.unit.converter(df,df.datadict,base_var.name,base_var.unit)

  df.display <- as.data.frame(df$year)
  
  for (i in 1:length(var.names)) {
    var <- variable.unit.converter(df,df.datadict,var.names[i],var.units[i])
    prop <- var/base_var
    df.display <- as.data.frame(cbind(df.display,prop))
  }
  
  colnames(df.display) <- c('year',var.names)
  df.display <- df.display %>% pivot_longer(cols=var.names)
  
  # myplot <- suppressMessages(plotly_build(plot_ly(data=df.display,
  #         x=~year,mode='lines+markers', split= ~name, color = ~name) %>%
  #   add_trace(y = ~value) %>%
  #   layout(legend=list(title=list(text='Variables')),
  #          xaxis = list(title = 'Year'),
  #          yaxis = list(title = paste(base_var.name, paste('(', base_var.unit,')',sep=''),sep=' ')))))
  # return(myplot)
  
  color_map <- setNames(c("blue", "orange", "green"), var.names.raw)
  plot_ly(data=df.display, 
          x=~year,
          y=~value,
          split=~name, 
          color=~name,
          colors=color_map,
          mode='lines+markers')
```


This section defines drop-down lists (Variable names, etc.)
```{r}
#Set Drop-Down lists
variables.list <- as.data.frame(colnames(df.orig))
colnames(variables.list)<-c("vars")
variables.list <- variables.list %>% 
  subset(!vars %in% c('X','year')) %>%
  filter(!grepl('hours', vars))
variables.list
```

```{r}
#Define Color Palette for Graphs
palette.main<-brewer.pal(n = 8, name = "Spectral")
ty.color<-palette.main[7]
ly.color<-palette.main[3]
lly.color<-palette.main[1]
```

```{r}
#Set column names for drop-down.
friendly_disease_list<-c(
  "All_Causes",
  "All_Natural_Causes",
  "Septicemia",
  "Cancer",
  "Diabetes",
  "Alzheimers",
  "Influenza_and_Pneumonia",
  "Chronic_Lowe_Respiratory_Diseases",
  "Other_Respiratory_Diseases",
  "Nephritis(Kidney_Problems)",
  "Not_Elsewhere_Classified",
  "Heart_Diseases",
  "Cerebrovascular(Stroke,Aneurysm)",
  "Covid_19_Multiple_COD",
  "Covid_19_Underlying_COD"
)
```

Below is the reactive (input.r) variables.  Comments are from butchered version.
```{r}
  input.r <- reactiveValues(
    df.display = df.orig,
    base_var.name = 'ground_beef',
    base_var.unit = 'pound',
    var.names = c('retail','financial'),
    var.units = c('hour','hour')
  )
    # df.ytd.1 = subset(df.main.1,MMWR.Year==update.year),
    # df.ty.1 = subset(df.main.1,MMWR.Year==last.year&ty_flag==TRUE),
    # df.ly.1 = subset(df.main.1,MMWR.Year==last.year),
    # df.lly.1 = subset(df.main.1,MMWR.Year==last.last.year),
    # df.ytd.2 = subset(df.main.2,MMWR.Year==update.year),
    # df.ty.2 = subset(df.main.2,MMWR.Year==last.year&ty_flag==TRUE),
    # df.ly.2 = subset(df.main.2,MMWR.Year==last.year),
    # df.lly.2 = subset(df.main.2,MMWR.Year==last.last.year),
    # current.year.total.1 = 1,
    # last.year.total.1 = 2,
    # last.last.year.total.1 = 3,
    # current.year.total.2 = 4,
    # last.year.total.2 = 5,
    # last.last.year.total.2 = 6,
    # region.1.r = "United States",
    # region.2.r = "United States",
    # cod_1.r = "Covid_19_Underlying_COD",
    # cod_2.r = "Covid_19_Multiple_COD",
    # current.year.total.1 = sum(subset(df.main.1,MMWR.Year==update.year)["disp"],na.rm=TRUE),
    # #Map
    # df.map = df.map,
    # df.map.active = subset(df.map,MMWR.Year==last.year & MMWR.Week==1),
    # map.year.r = last.year,
    # map.week.r = 1,
    # # last.year.total.1 = sum(df.ly.1["disp"],na.rm=TRUE),
    # # last.last.year.total.1 = sum(df.lly.1["disp"],na.rm=TRUE),
    # # current.year.total.2 = sum(df.ytd.1["disp"],na.rm=TRUE),
    # # last.year.total.2 = sum(df.ly.1["disp"],na.rm=TRUE),
    # # last.last.year.total.2 = sum(df.lly.1["disp"],na.rm=TRUE),
    # view.r = c(TRUE,FALSE,FALSE,TRUE,FALSE,FALSE), #Years to display: first 3 plot 1, second 3 plot 2.  TY/LY/LLY.
```
```{r}
# #Set main dataframes to filter from
df.display <- df.orig
```

Below is for server side observed elements.
regionSelector1 -> baseVarSelector
regionSelector2 -> REMOVED

Old
```{r}
  #Top Region Selector
  output$regionSelector <- renderUI({
    selectInput(
      inputId = "myRegion1",
      label = "Select a State, or 'United States' to view all:",
      choices = region.list
    )
  })

  #Reacts to UI element 'regionSelector1'
  observeEvent(input$myRegion1, {
    input.r$region.1.r <- input$myRegion1
    print(paste('Region 1 changed, active region is now:',input.r$region.1.r))
    input.r$df.ytd.1 <-   subset(input.r$df.main.1,Jurisdiction.of.Occurrence==input.r$region.1.r & MMWR.Year==update.year)
    input.r$df.ty.1 <-    subset(input.r$df.main.1,Jurisdiction.of.Occurrence==input.r$region.1.r & MMWR.Year==last.year&ty_flag==TRUE)
    input.r$df.ly.1 <-    subset(input.r$df.main.1,Jurisdiction.of.Occurrence==input.r$region.1.r & MMWR.Year==last.year)
    input.r$df.lly.1 <-   subset(input.r$df.main.1,Jurisdiction.of.Occurrence==input.r$region.1.r & MMWR.Year==last.last.year )
  }) 
```

New
```{r}
  #Variable Selector

  #Reacts to UI element 'baseVarSelector'
  observeEvent(input$baseVar1, {
    input.r$base.var.1.r <- input$baseVar1
    print(paste('Base Variables changed, active base variable is now:',input.r$region.1.r))
  }) 
```